An Extension of LTL with Rules\\ and its Application to Runtime Verification

### 2019-02-02 rule-ltl.tex
fixing problems

\section{Propositional Linear Temporal Logic}

@@

\section{Making propositional LTL more Expressive}

@@

@@

We present now an extension of LTL, which we call ELTL. It has the same
expressive power as  QLTL.
It has a form that puts it, conceptually, in between temporal
logic and automata: on the one hand, it is possible to simply represent
a pure PLTL formula, and on the other hand, one can represent a specification that resembles a finite automaton. As will be showed later, this extension is
is very natural for runtime verification.

@@

\begin{lemma} \label{fourone}
Let $\varphi$ be an ELTL formula as in~(\ref{ELTL}),
over the auxiliary propositions $A$ and basic propositions $B$ 
and a model $\sigma$ with states over the $B$, such that
$\sigma \models \psi$.
Then there is a unique
model $\sigma'$ such that $\sigma' |_A = \sigma$.

@@

%\klaus{I think this lemma needs to tie the trace to the formula. Doron: I do not understand this comment.}

@@

\begin{theorem}
\label{theo1}
The expressive power of 
ELTL is the same as QLTL.
\end{theorem}
\noindent {\bf Proof.}
Each ELTL formula $\eta$, as defined above,
is expressible in QLTL using the following formula:
\[   \Exists a_1 \ldots \Exists a_n  ( \varphi  \wedge   \Box \bigwedge_{1 \leq j \leq n} ( a_j \leftrightarrow \varphi_j))
\]
For the other direction, one can translate any QLTL property into an $\omega$ automaton\footnote{The past operators of a QLTL property do not change this fact, since the translation can be done first to second order monadic logic, which can be translated into an $\omega$-automaton}.
We assume that the property is given as a deterministic Muller automaton $M = (S, i , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $i \in S$ is the initial state, the set $\Sigma$ consists of the inputs, each represented as a Boolean combination of the variables $B$, $\Delta : S \times \Sigma \mapsto S$, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set, 
where an accepted input needs to pass infinitely often in exactly
one of the specified sets. To formulate an ELTL formula,
he auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.

@@

\section{Runtime Verification for PLTL and its extensions}
\label{LTLruntime}

@@

\section{First-Order Linear Temporal Logic FLTL}

@@

\section{Extending First-Order Temporal Logic}

@@

\section{A Runtime Verification Algorithm for EPFLTL}
\label{EPLTLRV}

>>>>

\section{Propositional LTL}

@@

\section{Extending Propositional LTL}

@@

\klaus{Are the last two lines above needed?}

@@

\subsubsection{Rule-based LTL}

We present now an extension of LTL, which we call ELTL. It has the same
expressive power as  QLTL.
It has a form that puts it, conceptually, in between temporal
logic and automata: on the one hand, it is possible to simply write
a pure LTL formula, and on the other hand, one can write a specification that resembles a finite automaton. As will be showed later, this extension is
is very natural for runtime verification.

@@

\begin{lemma} \label{fourone}
Let $\varphi$ be an ELTL formula as in~(\ref{ELTL}),
over the auxiliary propositions $A$ and basic propositions $B$, 
and let $\sigma$ be a model with states over the $B$, such that
$\sigma \models \psi$.
Then there is a unique
model $\sigma'$ such that $\sigma' |_A = \sigma$.

@@

\klaus{(1) is this lemma saying the right thing? (2) is the proof really a proof of that lemma, (3) 
should $\varphi$ be $\psi$, (4) now $\psi$ is not the whole spec (it is missing the rules). What do we call the whole thing? It is not a formula like the 
$\varphi$s and the $\psi$s. Do we need a special greek symbol for it?}

@@

\begin{theorem}
\label{theo1}
The expressive power of 
ELTL is the same as QLTL.
\end{theorem}
\noindent {\bf Proof.}
Each ELTL formula $\eta$, as defined above,
is expressible in QLTL using the following formula:
\[   \Exists a_1 \ldots \Exists a_n  ( \psi  \wedge   \Box \bigwedge_{1 \leq j \leq n} ( a_j \leftrightarrow \varphi_j))
\]
For the other direction, one can translate any QLTL property into an $\omega$ automaton\footnote{The past operators of a QLTL property do not change this fact, since the translation can be done first to second order monadic logic, which can be translated into an $\omega$-automaton}.
We assume that the property is given as a deterministic Muller automaton $M = (S, i , \Sigma, \Delta, {\cal F})$ (a B\"{u}chi automaton may not always be deterministic), where $S$ is the set of states, $i \in S$ is the initial state, the set $\Sigma$ consists of the inputs, each represented as a Boolean combination of the variables $B$, $\Delta : S \times \Sigma \mapsto S$, and ${\cal F} = \{ F_1 , \ldots , F_k \}$ is the accepting set, 
where an accepted input needs to pass infinitely often in exactly
one of the specified sets. To formulate an ELTL formula,
he auxiliary variables represent the states $S$ of $M$, i.e., $a_s$ represents the state $s \in S$.

@@

\section{RV for Propositional Past-Time LTL and its Extension}
\label{LTLruntime}

@@

\section{First-Order LTL}

@@

\section{Extending First-Order LTL}

@@

\section{RV for First-Order Past-Time LTL and its Extension}
\label{EPLTLRV}
